---
title: "2016 Election Prediction - Part 1"
author: junfeng
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

indent1 = '    '
indent2 = paste(rep(indent1, 2), collapse='')
indent3 = paste(rep(indent1, 3), collapse='')

library(knitr)
library(tidyverse)
library(ggmap)
library(maps)
library(Rtsne)
library(NbClust)
library(tree)
library(maptree)
library(class)
library(reshape2)
```

Predicting voter behavior is complicated for many reasons despite the tremendous effort in collecting, analyzing, and understanding many available datasets. 
For our final project, we will analyze the 2016 presidential election dataset, but, first, some background.

# Background

The presidential election in 2012 did not come as a surprise. Some correctly predicted the outcome of the election correctly including [Nate Silver](https://en.wikipedia.org/wiki/Nate_Silver), 
and [many speculated his approach](https://www.theguardian.com/science/grrlscientist/2012/nov/08/nate-sliver-predict-us-election).

Despite the success in 2012, the 2016 presidential election came as a 
[big surprise](https://fivethirtyeight.com/features/the-polls-missed-trump-we-asked-pollsters-why/) 
to many, and it was a clear example that even the current state-of-the-art technology can surprise us.

Answer the following questions in one to two paragraphs for each.

1. What makes voter behavior prediction (and thus election forecasting) a hard problem?
 it involves voting at the state level (residents in each state vote for the candidate they support, and the winning candidate gets all of the state's electoral votes). The polls are also arranged at both state and national level, so one way or another both need to be taken into account.
2. What was unique to Nate Silver's approach in 2012 that allowed him to achieve good predictions?
Instead of looking at the maximum probability, he looks at the full range of probabilities. 
3. What went wrong in 2016? What do you think should be done to make future predictions better?
Errors of all of  types(miss in same or different direction) added up to Tuesday’s result,All of them are important because all of them were present in this result.
i think they should be looking at a 4-point or so national miss – which as noted in the past by FiveThirtyEight is not an insane level of error, but it is real error and the public’s right to question polls is justified,and has some 
# Data

```{r data}
election.raw = read.csv("election.csv") %>% as.tbl
census_meta = read.csv("metadata.csv", sep = ";") %>% as.tbl
census = read.csv("census.csv") %>% as.tbl
census$CensusTract = as.factor(census$CensusTract)
```

## Election data

Following is the first few rows of the `election.raw` data:

```{r, echo=FALSE}
kable(election.raw %>% head)
```

The meaning of each column in `election.raw` is clear except `fips`. The accronym is short for [Federal Information Processing Standard](https://en.wikipedia.org/wiki/FIPS_county_code).

In our dataset, `fips` values denote the area (US, state, or county) that each row of data represent: i.e., some rows in `election.raw` are summary rows. These rows have `county` value of `NA`. There are two kinds of summary rows:

* Federal-level summary rows have `fips` value of `US`.
* State-level summary rows have names of each states as `fips` value.

## Census data

Following is the first few rows of the `census` data:

```{r, echo=FALSE}
kable(census %>% head)
```

### Census data: column metadata

Column information is given in `metadata`.

```{r, dependson=data, echo=FALSE}
kable(census_meta)
```

## Data wrangling
4. Remove summary rows from `election.raw` data: i.e.,

    * Federal-level summary into a `election_federal`.
    
    * State-level summary into a `election_state`.
    
    * Only county-level data is to be in `election`.

```{r num4}
election_federal<-election.raw %>% filter(fips=='US')
election_state<-election.raw %>% filter(is.na(county)) %>% filter(fips!='US')
election<-election.raw %>% filter(!is.na(county))
```
5. How many named presidential candidates were there in the 2016 election? Draw a bar chart of all votes received by each candidate
```{r num5}
votes=election_federal %>% group_by(candidate) %>%summarise(votes)
candidate1=data.frame(
  candidate=election_federal$candidate,
  vote_num=votes
)
length(candidate1$candidate)
barchart(candidate1$votes~candidate1$candidate,candidate1)
```
6. Create variables `county_winner` and `state_winner` by taking the candidate with the highest proportion of votes. 
  Hint: to create `county_winner`, start with `election`, group by `fips`, compute `total` votes, and `pct = votes/total`. 
  Then choose the highest row using `top_n` (variable `state_winner` is similar).
```{r num6}
total=election %>% group_by(fips) %>% summarise_each(funs(sum)/funs(n),votes)
```
    
## Visualization

Visualization is crucial for gaining insight and intuition during data mining. We will map our data onto maps.

The R package `ggplot2` can be used to draw maps. Consider the following code.

```{r, message=FALSE}
states = map_data("state")

ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  guides(fill=FALSE)  # color legend is unnecessary and takes too long
```

The variable `states` contain information to draw white polygons, and fill-colors are determined by `region`.

7. Draw county-level map by creating `counties = map_data("county")`. Color by county


#. Now color the map by the winning candidate for each state. 
  First, combine `states` variable and `state_winner` we created earlier using `left_join()`. 
  Note that `left_join()` needs to match up values of states to join the tables; however, they are in different formats: e.g. `AZ` vs. `arizona`.
  Before using `left_join()`, create a common column by creating a new column for `states` named
  `fips = state.abb[match(some_column, some_function(state.name))]`. 
  Replace `some_column` and `some_function` to complete creation of this new column. Then `left_join()`.
  Your figure will look similar to state_level [New York Times map](https://www.nytimes.com/elections/results/president).



#. The variable `county` does not have `fips` column. So we will create one by pooling information from `maps::county.fips`.
  Split the `polyname` column to `region` and `subregion`. Use `left_join()` combine `county.fips` into `county`. 
  Also, `left_join()` previously created variable `county_winner`. 
  Your figure will look similar to county-level [New York Times map](https://www.nytimes.com/elections/results/president).

  
#. Create a visualization of your choice using `census` data. Many exit polls noted that 
    [demographics played a big role in the election](https://fivethirtyeight.com/features/demographics-not-hacking-explain-the-election-results/).
    Use [this Washington Post article](https://www.washingtonpost.com/graphics/politics/2016-election/exit-polls/) 
    and [this R graph gallery](https://www.r-graph-gallery.com/) for ideas and inspiration.
    
#. The `census` data contains high resolution information (more fine-grained than county-level).  
    In this problem, we aggregate the information into county-level data by 
    computing `TotalPop`-weighted average of each attributes for each county. Create the following variables:
    
    * _Clean census data `census.del`_: 
      start with `census`, filter out any rows with missing values, 
      convert {`Men`, `Employed`, `Citizen`} attributes to a percentages (meta data seems to be inaccurate), 
      compute `Minority` attribute by combining {Hispanic, Black, Native, Asian, Pacific}, remove {`Walk`, `PublicWork`, `Construction`}.  
      _Many columns seem to be related, and, if a set that adds up to 100%, one column will be deleted._  
      

    * _Sub-county census data, `census.subct`_: 
      start with `census.del` from above, `group_by()` two attributes {`State`, `County`}, 
      use `add_tally()` to compute `CountyTotal`. Also, compute the weight by `TotalPop/CountyTotal`.
    

    * _County census data, `census.ct`_: 
      start with `census.subct`, use `summarize_at()` to compute weighted sum
    

    * _Print few rows of `census.ct`_: 
    

## Dimensionality reduction

12. Run PCA for both county & sub-county level data. Save the first two principle components PC1 and PC2 into a two-column data frame, call it `ct.pc2` and `subct.pc2`, respectively.


#. Consider the principal component loadings PC1 and PC2. Which attributes do the most prominent loadings correspond to?
 
#. Create a scatter plot of two principal components, then color the points by the attributes corresponding to the prominent loadings.  
      

